---
title: "PM566 Lab 4"
author: "Misha Khan"
date: "9/14/2022"
output: html_document
---
```{r install-libraries}
library(lubridate)
library(tidyverse)
library(data.table)

```


## 1. Load data
```{r load-data, cache = TRUE}
#cache = TRUE saves the data
if (!file.exists("../lab03/met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
  )
met <- data.table::fread("met_all.gz")
```

## 2. Prepare the data

Remove temperatures less than -17C abd cgabge ekev 9999 to missing value code
```{r remove-low-temp}
met <- met[temp > -17][elev == 9999.0, elev := NA]
```

Make sure there are no missing data in the key variables coded as 9999, 999, etc
```{r}
#just another way you can do it separate
#met[met$elev == 9999.0] <- NA
#summary(met$elev) #check NAs
```

Generate a date variable using the functions as.Date()
```{r create-ymd, cache = TRUE}
#paste YMD and sep is the - in between
#[ , new col]
met <- met[ , ymd := as.Date(paste(year, month, day, sep = "-"))]
```

Using the data.table::week function, keep the observations of the first week of the month.
```{r}
met[ , table(week(ymd))] #frequency of each week 31-35
met$ymd[1:5] #August weeks
met<- met[week(ymd) == 31] #subsetting data to that week
nrow(met) #reduced dataset to 297259 rows (first week of August)
```

Check no 9999s in other important variables
```{r check-max}
met[ , .(
  temp = max(temp, na.rm = T),
  rh = max(rh, na.rm = T),
  wind.sp = max(wind.sp, na.rm = T),
  vis.dist = max(vis.dist, na.rm = T),
  dew.point = max(dew.point, na.rm = T),
  lat = max(lat, na.rm = T),
  lon = max(lon, na.rm = T),
  elev = max(elev, na.rm = T)
)]
```

Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.
```{r check-mean}
met_avg <- met[ , .(
  temp = mean(temp, na.rm = T),
  rh = mean(rh, na.rm = T),
  wind.sp = mean(wind.sp, na.rm = T),
  vis.dist = mean(vis.dist, na.rm = T),
  dew.point = mean(dew.point, na.rm = T),
  lat = mean(lat, na.rm = T),
  lon = mean(lon, na.rm = T),
  elev = mean(elev, na.rm = T)
), by = "USAFID"]
```

Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees
```{r}
#:= means assign
#first condition, assign NE NW SW, otherwise SE
met_avg[, region := fifelse(lon > -98 & lat > 39.71, "NE", 
                fifelse(lon < -98 & lat > 39.71, "NW",
                fifelse(lon < -98 & lat <= 39.71, "SW",
                "SE")))
    ]

table(met_avg$region)
```

Create a categorical variable for elevation as in the lecture slides
```{r}
met_avg[, elev_cat := fifelse(elev > 252, "high", "low")]
```


## 3. Use geom_violin to examine the wind speed and dew point temperature by region
Use facets
Set the x aesthetic to 1
Make sure to deal with NA category
Describe what you observe in the graph
```{r violin-dewpoint}
met_avg[!is.na(region) & !is.na(dew.point)] %>% 
  ggplot() + 
  geom_violin(mapping = aes(y = dew.point, x = 1, color=region, fill = region)) + 
  facet_wrap(~ region, nrow = 1)
```


The highest dew point temperatures are in the southeast (blue).

```{r violin-windspeed}
met_avg[!is.na(region) & !is.na(wind.sp)] %>% 
  ggplot() + 
  geom_violin(mapping = aes(y = wind.sp, x = 1, color=region, fill = region)) + 
  facet_wrap(~ region, nrow = 1)
```
Windspeeds are fairly low across all regions

## 4. Use geom_point with geom_smooth to examine the association between dew point temperature and wind speed by region
Colour points by region
Make sure to deal with NA category
Fit a linear regression line by region
Describe what you observe in the graph
```{r scatterplot-dewpoint}
met_avg[!is.na(region) & !is.na(wind.sp)] %>% 
  ggplot(mapping = aes(x = wind.sp, y= dew.point)) + 
  geom_point(mapping = aes(color = region)) + 
  geom_smooth(method = lm, mapping = aes(linetype = region)) +
  facet_wrap(~ region, nrow = 2) #facet wrap makes separate graphs
```













