---
title: "PM566 Homework 2"
author: "Misha Khan"
date: "2022-10-06"
output: github_document
always_allow_html: true
---
For this assignment, we will be analyzing data from USC’s Children’s Health Study. The learning objectives are to conduct data wrangling and visualize the data with key questions in mind.
```{r install-libraries}
library(lubridate)
library(tidyverse)
library(data.table)  
library(dtplyr)
library(dplyr)
library(readr)
library(Hmisc)
library(skimr) 
library(leaflet)
library(webshot)
library(psych)
```

# Data Wrangling
Two data sets: individual and regional CHS
**individual - personal and health characteristics of children in 12 communities in Socal**
**regional - air quality at a community level**
### Merge data sets using location variable

```{r read-data}
indv <- read.csv("chs_individual.csv")
reg <- read.csv("chs_regional.csv")
```

Merged data set: chs
```{r}
#Location variable = "townname"
chs <- merge(
  # Data
  x     = indv,      
  y     = reg, 
  # List of variables to match
  by  = "townname",
  # Which obs to keep?
  all.x = TRUE,      
  all.y = FALSE
  )

```

## 1. After merging the data, make sure you don’t have any duplicates by counting the number of rows. Make sure it matches.
```{r}
skimr::skim(chs)
```
indv data set: 1200 rows, 23 columns
reg data set: 12 rows, 27 columns
chs data set: 1200 rows, 49 columns

When the data sets are merged, the number of rows still stays the same but the columns increases (23 + 27 - 1 = 49 total cols).
There is no duplicate number of rows, everything matches (nothing lost or duplicated).

### In the case of missing values, impute data using the average within the variables “male” and “hispanic.” 
```{r}
#Output the number of columns with number of missing values
which(colSums(is.na(chs)) > 0)
```
```{r}
#chs <- data.table(chs)
#chs[, bmi_imp := fcoalesce(bmi, mean(bmi, na.rm = TRUE)), by = .(male, hispanic)]
#chs[ , fev_imp := fcoalesce(fev, mean(fev, na.rm = TRUE)), by = .(male, hispanic)]
```

## 2. Create a new categorical variable named “obesity_level” using the BMI measurement (underweight BMI<14; normal BMI 14-22; overweight BMI 22-24; obese BMI>24). 
```{r}
chs <-
  chs %>%
  mutate(obesity_level = case_when(chs$bmi < 14  ~ 'Underweight', 
                              chs$bmi >= 14 & chs$bmi < 22  ~ 'Normal', 
                              chs$bmi >= 22 & chs$bmi < 24  ~ 'Overweight', 
                              chs$bmi >= 24  ~ 'Obese'))
```

### To make sure the variable is rightly coded, create a summary table that contains the minimum BMI, maximum BMI, and the total number of observations per category.
```{r}
chs %>%
  group_by(obesity_level) %>%
  summarise(
    min_bmi  = min(bmi), 
    max_bmi  = max(bmi),
    mean_bmi = mean(bmi),
    n_bmi    = n()) %>%
knitr::kable()
```

## 3. Create another categorical variable named “smoke_gas_exposure” that summarizes “Second Hand Smoke” and “Gas Stove.” The variable should have four categories in total
```{r}
chs <-
  chs %>%
  mutate(smoke_gas_exposure = case_when(chs$smoke == 0 & chs$gasstove == 0 ~ 'No exposure',
                                        chs$smoke == 1 & chs$gasstove == 1 ~ 'Smoke & gas exposure',
                                        chs$smoke == 0 & chs$gasstove == 1 ~ 'Gas exposure',
                                        chs$smoke == 1 & chs$gasstove == 0 ~ 'Smoke exposure'))
table(chs$smoke_gas_exposure)
```


## Create four summary tables showing the average (or proportion, if binary) and sd of “Forced expiratory volume in 1 second (ml)” and asthma indicator by town, sex, obesity level, and “smoke_gas_exposure.”
```{r}
chs %>%
  group_by(townname) %>%
  summarise(
    n    = n(),
    mean_fev  = mean(fev), 
    sd_fev  = sd(fev),
    mean_fev = mean(fev))
    #prop_asthma = sum(asthma)/n()*100) %>%
  #distinct(townname, .keep_all = T) %>%
#knitr::kable()
```
# EDA
Primary questions of interest
**What is the association between BMI and FEV?**
**What is the association between smoke and gas exposure and FEV?**
**What is the association between PM2.5 and FEV?**

## Checklist
```{r}
#Check the dimensions and headers and footers of the data
dim(chs)
```

```{r}
#Check the variable types in the data
str(chs)
```

```{r}
#Conduct some summary statistics to answer the initial question
summary(chs$bmi)
summary(chs$fev)
summary(chs$smoke)
summary(chs$gasstove)
summary(chs$pm2_5_fr)
```

## 1. Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.

```{r}
chs %>%
  ggplot(aes(x = bmi, y = fev, group = townname)) +
  geom_point(mapping = aes(x = bmi, y = fev, color = townname)) +
  geom_smooth(se = T, method = "lm") +
  scale_color_brewer(palette = "Paired") +
  facet_wrap( ~ townname, nrow= 3) +
  ggtitle("BMI vs FEV per Town") +
  xlab("BMI") +
  ylab("FEV (mL)")
```
All the towns show a positive linear relationship with BMI and FEV meaning as BMI increases,
FEV also increases. Towns like Lake Elsinore, Lompoc, Long Beach, San Dimas, Santa Maria, and Upland
have a fairly steeper slope just at a glace. Additional statistical analysis will need to be conducted
to confirm.

# 2. Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.
```{r}
chs %>%
  ggplot(mapping = aes(x = fev, fill = obesity_level)) + 
  geom_histogram(position = "identity", alpha = 0.7, binwidth = 100, color = "black") +
  scale_fill_brewer(palette = "Dark2") +
  ggtitle("FEV per Obesity Level Distribution") +
  xlab("FEV (mL)") +
  ylab("Count")
```
Here is the distribution breakdown of FEV per obesity levels. Normal BMI level looks like
to have a normal distribution. The majority of individuals with normal BMI level have an FEV
around ~2000 mL on average. We can see that individuals with a underweight BMI level have a lower
FEV while those with obese BMI level have a higher FEV.

```{r}
chs %>%
  ggplot(mapping = aes(x = fev, fill = smoke_gas_exposure)) + 
  geom_histogram(position = "identity", alpha = 0.7, binwidth = 100, color = "black") +
  scale_fill_brewer(palette = "Set1") +
  ggtitle("FEV per Smoke Gas Exposure Distribution") +
  xlab("FEV (mL)") +
  ylab("Count")
```
Here is a distribution breakdown of FEV per smoke gas exposure. The majority of individuals
fall into the gas exposure category. The gas exposure distribution lies around the middle ~2000 mL.
The rest of the other categories also lie around the middle but have lower count of individuals. 

## 3. Barchart of BMI by smoke/gas exposure.
```{r}
chs %>%
  ggplot(mapping = aes(x = obesity_level, fill = factor(smoke_gas_exposure))) + 
  geom_bar(position = "dodge") +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("BMI per Smoke Gas Exposure") +
  xlab("BMI") +
  ylab("Obesity Level") +
  labs(fill = "Exposure")
```
Here is a barchart of smoke and gas exposure per obesity level. 
Individuals with a normal BMI level have the highest exposure to gas (over 450).
Across all BMI categories, gas exposure is the highest.

## 4. Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.
```{r}
chs %>%
    ggplot(mapping = aes(x = obesity_level, y = fev, color = obesity_level)) + 
    stat_summary(fun.data = mean_sdl, geom = "pointrange") +
    stat_summary(fun.data = mean_sdl, geom = "errorbar")+
  labs (title = "Statisitical Summary of FEV by BMI")
```

ADD INTERPRETATION


```{r}
chs %>%
    ggplot(mapping = aes(x = smoke_gas_exposure, y = fev, color = smoke_gas_exposure)) + 
    stat_summary(fun.data = mean_sdl, geom = "pointrange") +
    stat_summary(fun.data = mean_sdl, geom = "errorbar")+
  labs (title = "Statisitical Summary of FEV by Exposure")
```

ADD INTERPRETATION

## 5. A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.
```{r}
pm.pal <- colorFactor(c('red', 'yellow', 'blue', 'green'), domain = chs$pm25_mass)
```


```{r}
#chs[!is.na(pm25_mass)]%>%
  pm_map <- 
  leaflet(chs) %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addCircles(
    lat = ~lat, 
    lng = ~lon,
    label = ~paste0(chs$pm25_mass),
    color = ~ pm.pal(chs$pm25_mass),
    opacity = 0.5,
    fillOpacity = 1,
    radius = 500
    ) %>%
  addLegend('bottomleft', 
            pal = pm.pal, 
            values = chs$pm25_mass,
            title = 'PM 2.5 Mass', 
            opacity = 1)
pm_map
```
We can see from the leaflet plot that the higher concentration of PM 2.5 mass are in more centrally located
in Southern California (Ontario, Riverside). As we move away from that area, PM 2.5 mass decreases.
Areas like San Luis Obispo, Santa Maria, Santa Barbara, and San Diego show lower levels of PM 2.5 mass concentration.

## 6. Choose a visualization to examine whether PM2.5 mass is associated with FEV.
```{r}
chs %>%
    ggplot(mapping = aes(x = pm25_mass, y = fev, color = fev)) + 
    geom_point(mapping = aes(x = pm25_mass, y = fev)) +
    scale_color_gradient(low = "blue", high = "red") +
    geom_jitter() +
    geom_smooth(method = lm, se = F, col = "black")
```
Above is a graph of the relationship between PM 2.5 mass and FEV.
We see three separate ranges of PM 2.5 mass: 10, 20, 30.
As the PM 2.5 mass increases, the black line shows a slight decrease in FEV.
PM 2.5 mass of 10 has the highest FEV points while PM 2.5 mass 30 is slighty lower.





```{r}
#Group the columns with missing values (integer and numeric cols)
fix_na <- c("agepft", 
            "height", 
            "weight", 
            "bmi", 
            "asthma", 
            "father_asthma", 
            "mother_asthma", 
            "wheeze", 
            "hayfever",
            "allergy",
            "educ_parent",
            "smoke", 
            "gasstove",
            "fev",
            "fvc",
            "mmef",
            "no_24hr",
            "pm2_5_fr")

sum(is.na(chs[fix_na]))

```

