---
title: "PM566 Lab 4"
author: "Misha Khan"
date: "9/14/2022"
output: github_document
always_allow_html: true
---
```{r install-libraries}
library(lubridate)
library(tidyverse)
library(data.table)
library(leaflet)
```
   
 
## 1. Load data
```{r load-data, cache = TRUE}
#cache = TRUE saves the data
if (!file.exists("../lab04/met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
  )
met <- data.table::fread("met_all.gz")
```

## 2. Prepare the data

Remove temperatures less than -17C abd cgabge ekev 9999 to missing value code
```{r remove-low-temp}
met <- met[temp > -17][elev == 9999.0, elev := NA]
```

Make sure there are no missing data in the key variables coded as 9999, 999, etc
```{r}
#just another way you can do it separate
#met[met$elev == 9999.0] <- NA
#summary(met$elev) #check NAs
```

Generate a date variable using the functions as.Date()
```{r create-ymd, cache = TRUE}
#paste YMD and sep is the - in between
#[ , new col]
met <- met[ , ymd := as.Date(paste(year, month, day, sep = "-"))]
```

Using the data.table::week function, keep the observations of the first week of the month.
```{r}
met[ , table(week(ymd))] #frequency of each week 31-35
met$ymd[1:5] #August weeks
met<- met[week(ymd) == 31] #subsetting data to that week
nrow(met) #reduced dataset to 297259 rows (first week of August)
```

Check no 9999s in other important variables
```{r check-max}
met[ , .(
  temp = max(temp, na.rm = T),
  rh = max(rh, na.rm = T),
  wind.sp = max(wind.sp, na.rm = T),
  vis.dist = max(vis.dist, na.rm = T),
  dew.point = max(dew.point, na.rm = T),
  lat = max(lat, na.rm = T),
  lon = max(lon, na.rm = T),
  elev = max(elev, na.rm = T)
)]
```

Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.
```{r check-mean}
met_avg <- met[ , .(
  temp = mean(temp, na.rm = T),
  rh = mean(rh, na.rm = T),
  wind.sp = mean(wind.sp, na.rm = T),
  vis.dist = mean(vis.dist, na.rm = T),
  dew.point = mean(dew.point, na.rm = T),
  lat = mean(lat, na.rm = T),
  lon = mean(lon, na.rm = T),
  elev = mean(elev, na.rm = T)
), by = "USAFID"]
```

Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees
```{r}
#:= means assign
#first condition, assign NE NW SW, otherwise SE
met_avg[, region := fifelse(lon > -98 & lat > 39.71, "NE", 
                fifelse(lon < -98 & lat > 39.71, "NW",
                fifelse(lon < -98 & lat <= 39.71, "SW",
                "SE")))
    ]

table(met_avg$region)
```

Create a categorical variable for elevation as in the lecture slides
```{r}
met_avg[, elev_cat := fifelse(elev > 252, "high", "low")]
```


## 3. Use geom_violin to examine the wind speed and dew point temperature by region
Use facets
Set the x aesthetic to 1
Make sure to deal with NA category
Describe what you observe in the graph
```{r violin-dewpoint}
met_avg[!is.na(region) & !is.na(dew.point)] %>% 
  ggplot() + 
  geom_violin(mapping = aes(y = dew.point, x = 1, color=region, fill = region)) + 
  facet_wrap(~ region, nrow = 1)
```


The highest dew point temperatures are in the southeast (blue).

```{r violin-windspeed}
met_avg[!is.na(region) & !is.na(wind.sp)] %>% 
  ggplot() + 
  geom_violin(mapping = aes(y = wind.sp, x = 1, color=region, fill = region)) + 
  facet_wrap(~ region, nrow = 1)
```
Windspeeds are fairly low across all regions

## 4. Use geom_point with geom_smooth to examine the association between dew point temperature and wind speed by region
Colour points by region
Make sure to deal with NA category
Fit a linear regression line by region
Describe what you observe in the graph
```{r scatterplot-dewpoint}
met_avg[!is.na(region) & !is.na(wind.sp)] %>% 
  ggplot(mapping = aes(x = wind.sp, y= dew.point)) + 
  geom_point(mapping = aes(color = region)) + 
  geom_smooth(method = lm, mapping = aes(linetype = region)) +
  facet_wrap(~ region, nrow = 2) #facet wrap makes separate graphs
```
NE and SE are faily low in wind speed but high in dew point. NW and SW are more scattered across dew point and low in wind speed.

##5. Use geom_bar to create barplots of the weather stations by elevation category coloured by region
Bars by elevation category using position="dodge"
Change colours from the default. Colour by region using scale_fill_brewer see this
Create nice labels on axes and add a title
Describe what you observe in the graph
Make sure to deal with NA

```{r}
met_avg[!is.na(elev & elev != "clear")] %>%
  ggplot() + 
  geom_bar(mapping = aes(x = region, fill = wind.sp), position = "dodge")  + scale_fill_brewer(palette = "Dark2") +
  labs(title = "Barplot of Weather Stations by Elevation")
```


##6. Use stat_summary to examine mean dew point and wind speed by region with standard deviation error bars
Make sure to remove NA
Use fun.data=“mean_sdl” in stat_summary
Add another layer of stats_summary but change the geom to "errorbar" (see the help).
Describe the graph and what you observe
```{r}
met_avg[!is.na(dew.point)] %>%
  ggplot(mapping = aes(x = region, y = dew.point)) + 
    stat_summary(fun.data = mean_sdl, geom = "pointrange") +
    stat_summary(fun.data = mean_sdl, geom = "errorbar")

#met[region =="NE, min(deew.point, na.rm = T)] #checking the sum stats
```
Dew point has a wider range for NW and SW regions but smaller range for NE and SE regions.

##7. Make a map showing the spatial trend in relative h in the US
Make sure to remove NA
Use leaflet()
Make a colour palette with custom colours
Use addMarkers to include the top 10 places in relative h (hint: this will be useful rank(-rh) <= 10)
Add a legend
Describe trend in RH across the US

Use addMarkers to include top 10 places in relative
```{r}
top10rh <- met_avg[rank(-rh) <= 10]
```

```{r}
#Generating color palette
rh.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=met_avg$rh)
rh.pal

rhmap <- leaflet(met_avg) %>% 
  # The looks of the Map
  addProviderTiles('CartoDB.Positron') %>% 
  # Some circles
  addCircles(
    lat = ~lat, lng=~lon,
                                                  # HERE IS OUR PAL!
    label = ~paste0(rh), color = ~ rh.pal(rh),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=rh.pal, values=met_avg$rh,
          title='Relative Humid.', opacity=1)
rhmap
```
Humidity is higher on the east coast compared to west coast. 

