---
title: "lab03"
author: "Misha Khan"
date: "`r format(Sys.Date())`"
output: github_document
always_allow_html: true
---

## Step 1. Read in the data
First download and then read in with data.table:fread()

```{r}
#if file does not exist in my dir, then download it (avoid downloading mult times)
if (!file.exists("met_all.gz")){
 download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", url = "met_all.gz",method="libcurl", timeout = 60) 
}

met <- data.table::fread("met_all.gz") 
```
## Step 2. Check the dimensions, headers, footers. How many columns, rows are there?
```{r}
dim(met)
head(met)
tail(met)
```
## Step 3. Take a look at the variables.
```{r}
str(met)
```
## Step 4. Take a closer look at the key variables.
```{r}
table(met$year)
table(met$day)
table(met$hour)
```
```{r}
summary(met$temp)
summary(met$elev)
summary(met$wind.sp)
```
Replace the elevations with 9999 as NA
```{r}
met[met$elev==9999.0] <- NA
summary(met$elev)
```

The weather station with the highest elevation is at `r max(met$elev, na.rm = T)` meters.

We also have the issue of the minimum temperature being -40C, so we should remove those observations.
```{r}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```
We again notice that there is a `r min(met$temp, na.rm = T)`C temperature reading that seems suspicious.

## Step 5. Check the data against an external data source.
```{r}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```
## Step 6. Calculate summary statistics

Select the weather station with maximum elevation
```{r}
#Subset of data is created
elev <- met[elev==max(elev)]
summary(elev)
```

```{r}
#Wind speeds at the highest elevation
met[elev == max(elev, na.rm = T), summary(wind.sp)]
#Wind speeds are between 0-21
```
```{r}
met[elev == max(elev, na.rm = T), summary(temp)]
#Temps are between 1-15 
```

Look at the correlation between temperature and wind speed; look at the correlation between temperature and wind speed with hour and day of the month
```{r}
#Use everything includes NA, use complete does not include NA values
met[elev == max(elev, na.rm = T), .(
  temp_wind     = cor(temp, wind.sp, use = "complete"),
  temp_hour     = cor(temp, hour, use = "complete"),
  wind.sp_day   = cor(wind.sp, day, use = "complete"),
  wind.sp_hour  = cor(wind.sp, hour, use = "complete")   
)]
```
#Step 7. Explanatory graphs
```{r}
hist(met$elev, breaks = 100)
hist(met$wind.sp)
hist(met$temp)
```
Look at where the weather station with highest elevation is located
```{r}
if(!require(leaflet)) {
  install.packages("leaflet")
  library(leaflet)
}

elev <- met[elev == max(elev)]
```


```{r}
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```

```{r}
library(lubridate)
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)
```
```{r}
elev <- elev[order(date)]
head(elev)
```

```{r}
plot(elev$date, elev$temp, type='l')
```
```{r}
plot(elev$date, elev$wind.sp, type='l')
```


