---
title: "lab06"
author: "Misha Khan"
date: "2022-09-28"
output: github_document
always_allow_html: true
---
```{r}
library(tidytext)
library(tidyverse)
library(data.table)
library(dplyr)
library(dtplyr)
library(ggplot2)
library(forcats)
```

```{r load-data, cache = TRUE}
#cache = TRUE saves the data
if (!file.exists("mtsamples.csv"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/00_mtsamples/mtsamples.csv",
    destfile = "mtsamples.csv",
    method   = "libcurl",
    timeout  = 60
  )
mts <- read.csv("mtsamples.csv")
dim(mts) 
```
```{r}
#this is a dataframe, not ideal for dplyr -> change to tibble
mts <- as_tibble(mts)
```

# Question 1: What specialties do we have?
We can use count() from dplyr to figure out how many different categories do we have? Are these catagories related? overlapping? evenly distributed?
```{r}
specialties <- mts %>%
  count(medical_specialty)

specialties %>%
  arrange(desc(n))%>%
knitr::kable()
 
```

There are `nrows(specialties)` medical specialties.

```{r barplot-of-specialties-count}
specialties %>%
  top_n(10) %>%
  ggplot(aes(x = n, y = fct_reorder(medical_specialty, n))) +
  geom_col()
  
```

Distribution is not uniform.

# Question 2
Tokenize the the words in the transcription column
Count the number of times each token appears
Visualize the top 20 most frequent words
Explain what we see from this result. Does it makes sense? What insights (if any) do we get?
```{r}
mts %>%
  unnest_tokens(word, transcription) %>%
  count(word, sort = TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```

There are a lot of stopwords here, non specific to medical text.
We do see "patient", phew!

# Question 3
Redo Q2 visualization but remove stopwords before
Bonus points if you remove numbers as well
What do we see know that we have removed stop words? Does it give us a better idea of what the text is about?
```{r}
mts %>%
  unnest_tokens(word, transcription) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = c("word")) %>%
 #use regex to filter our numbers [0-9] +$ means incl any number larger than 10
  filter(!grepl(pattern = "^[0-9]+$", x = word)) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```
Removing the stop words and numbers gives us a much better idea.

# Question 4
Repeat Q2 but this time tokenize into bi-grams. how does the result change if you look at tri-grams?
```{r}
mts %>%
  unnest_ngrams(bigram, transcription, n = 2) %>%
  count(bigram, sort = TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(bigram, n))) +
  geom_col()
```
```{r}
mts %>%
  unnest_ngrams(trigram, transcription, n = 3) %>%
  count(trigram, sort = TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(trigram, n))) +
  geom_col()
```
Trigrams found more medical word groups than bigrams

# Question 5
Using the results you got from questions 4. Pick a word and count the words that appears after and before it.
```{r}
ptbigram <-
  mts %>%
  unnest_ngrams(bigram, transcription, n = 2) %>%
  separate(bigram, into = c("word1", "word2"), sep = " ") %>%
  select(word1, word2) %>%
  filter(word1 == "patient" | word2 == "patient")

```

Words appearing before patient: 
```{r}
ptbigram %>%
  filter(word2 == "patient") %>%
  count(word1, sort = T) %>%
  anti_join(stop_words, by = c('word1' = 'word')) %>%
  top_n(10) %>%
knitr::kable()
```

Words appearing after patient: 
```{r}
ptbigram %>%
  filter(word1 == "patient") %>%
  count(word2, sort = T) %>%
  anti_join(stop_words, by = c('word2' = 'word')) %>%
  top_n(10) %>%
knitr::kable()
```
# Question 6
Which words are most used in each of the specialties. you can use group_by() and top_n() from dplyr to have the calculations be done within each specialty. Remember to remove stopwords. How about the most 5 used words?
```{r}
mts %>%
  unnest_tokens(word, transcription) %>%
  group_by(medical_specialty) %>%
  count(word, sort = TRUE) %>%
  filter(!(word %in% stop_words$word) & !grepl(pattern = "^[0-9]+$", x = word)) %>%
  top_n(5, n) %>%
  arrange(medical_specialty, desc(n))%>%
  knitr::kable()
```


